\begin{displayquote}
"If people do not believe that mathematics is simple, it is only because they do not realize how complicated life is." 
\end{displayquote}

\begin{flushright}
\textit{John von Neumann}
\end{flushright}


\newpage
\section*{Acknowledgments}

No man is an island.

While a single name appears on the cover of this document, it should not be implied that anything was, or could be done, alone. Everyone is the product of their environments and I've had the esteemed privilege to be assisted and inspired by many wonderful people. 

So my great appreciation to Ewa Podkolinska, Jerzy Podkolinski, Alina Kroeker, Raymond Kroeker, Tom Ruette, Wannes Meert, Mathias Verbeke, Gregory Schiano, Jan Aerts, Gerda Claeskens, Luc De Raedt, Roel Braekers, Geert Verbeke, Eric Schmitt, Rui Barbosa, Stijn Van Weezel, Akshat Dwivedi, Irzam Hardiansyah, Zhu Meng, Sytze Elzinga, Justin Fischedick and most importantly Rianne Hartemink.

Thank you for your support and motivation. 

\begin{flushright}
\textit{Richard Podkolinski}
\\
2016 
\\
Den Haag
\\
The Netherlands
\end{flushright}
\\

\section*{Intended Audience}

Good research is the kind that is capable of taking a complex topic making it accessible to a broader audience. In an effort to adhere to that standard, the intended audience of this thesis is not a pure mathematician, rather a general practitioner in the area of predictive maintenance or reliability analysis. Emphasized is the development of understanding and intuition for the underlying models and how to handle edge cases. 

While great effort is undertaken to be as accessible as possible, a certain degree of mathematical development is required to understand this text. Recommended is an understanding of the calculus of probabilities at the level of Blitzstein and Huang\cite{Blitzstein2014} as well as a first course on Bayesian linear models at the level of McElreath\cite{McElreath2016}. A basic understanding of how basic linear models are fit and their performance evaluated is required.

To ensure access to rigorous results, an extensive list of references and resources are made available at the end of each chapter, dubbed "Extended Reading". These resources elucidate many of the technical aspects of the modeling process and mathematical foundations which are beyond the scope of this text.